Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-02-11 VegOzone.tex
Start to write about latest results, remove obsolete text.

A comprehensive list of model parameters used as starting point in this study can be found in Supplement~Table~S1. These default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual}. Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each PFT are estimated as follows. Evergreen needle leaf trees are photosynthetic active already at low temperatures \citep{TB:Kolari2007}. This leads necessarily to an increase in observable net photosynthesis. Hence, we use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine onset and end of growing season. Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}). By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{day~of~year}$ for 2018 and day 106 in 2019. $A_\text{end}$ amounts to day 261 and 274, respectively, and shall be used for all vegetation alike. {\bf TODO: Compare with latitudinal model?} The resulting growing season for Norway spruce in 2019 was one month longer then in 2018. For estimating $A_\text{start}$ of birch tree, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}). We find 129 and 130, respectively {\bf Accuracy?}. For perennial grassland, we assume a latency period of about $1\,\unit{month}$ after snow melt. Data for first snow-free day is taken from the weather station at Øvre Neiden in Sør-Varanger. In both years, the first snow-free day incidents with the estimated date of bud burst. As leafs of birches in northern Fennoscandia \emph{Betula nana} {\bf (TODO: Right name?)} are smaller than southern specimen, we have measured their dimension for the purpose of this study and find $(3.0\pm 0.5)\,\unit{cm}$. \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
%To localize leaf area index (LAI), we roughly characterize land use for the $1\times 1\,\unit{km}$ patch centered at Svanhovd from areal pictures at kartverket.no: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban. We normalize LAI to deciduous trees based on standard $\mathrm{DO_3SE}$ parameterization: $f_\text{con}=5/4\cdot f_\text{dec}$, $f_\text{grass}=7/8\cdot f_\text{dec}$. With these, we re-scale the average maximum LAI from MODIS ($\left<f\right> = 2.7$). This results in 3.75, 3, and 2.625 for Norway spruce, birch, and grassland, respectively.

@@

\begin{figure*}[t]
  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3}
  \caption{$\mathrm{POD_y}$ modeling results from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October. A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen. \chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$. Due to non linearity in stomatal conductance, the climatological year represents rather an optimal year then a true climatology. (a)--(c) Birch; (d)--(f) Norway spruce; (g)--(i) perennial grassland. From left to right: 2018, 2019, and climatological year.}
  \label{fig:pody}
\end{figure*}

>>>>

A comprehensive list of model parameters used as starting point in this study can be found in Supplement~Table~S1. These default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual}. Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each PFT are estimated as follows. Evergreen needle leaf trees are photosynthetic active already at low temperatures \citep{TB:Kolari2007}. This leads necessarily to an increase in observable net photosynthesis. Hence, we use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine onset and end of growing season. Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}). By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{day~of~year}$ for 2018 and day 106 in 2019. $A_\text{end}$ amounts to day 261 and 274, respectively, and shall be used for all vegetation alike. {\bf TODO: Compare with laitude model? Birch: 100--307} The resulting growing season for Norway spruce in 2019 was one month longer then in 2018. For estimating $A_\text{start}$ of birch tree, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}). We find 129 and 130, respectively {\bf Accuracy?}. For perennial grassland, we assume a latency period of about $1\,\unit{month}$ after snow melt. Data for first snow-free day is taken from the weather station at Øvre Neiden in Sør-Varanger. In both years, the first snow-free day incidents with the estimated date of bud burst. As leafs of birches in northern Fennoscandia \emph{Betula nana} {\bf (TODO: Right name?)} are smaller than southern specimen, we have measured their dimension for the purpose of this study and find $(3.0\pm 0.5)\,\unit{cm}$. \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
%To localize leaf area index (LAI), we roughly characterize land use for the $1\times 1\,\unit{km}$ patch centered at Svanhovd from areal pictures at kartverket.no: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban. We normalize LAI to deciduous trees based on standard $\mathrm{DO_3SE}$ parameterization: $f_\text{con}=5/4\cdot f_\text{dec}$, $f_\text{grass}=7/8\cdot f_\text{dec}$. With these, we re-scale the average maximum LAI from MODIS ($\left<f\right> = 2.7$). This results in 3.75, 3, and 2.625 for Norway spruce, birch, and grassland, respectively.

@@

%\begin{figure*}[t]
%  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3}
%  \caption{$\mathrm{POD_y}$ modeling results from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October. A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen. \chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$. Due to non linearity in stomatal conductance, the climatological year represents rather an optimal year then a true climatology. (a)--(c) Birch; (d)--(f) Norway spruce; (g)--(i) perennial grassland. From left to right: 2018, 2019, and climatological year.}
%  \label{fig:pody}
%\end{figure*}
